Predicting Kamala Harris's Victory in the 2024 US Election

### 2024-10-25 paper/paper.qmd
Update model diagnosis section.

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Xuanle Zhou
  - Yongqi Liu
  - Yuxuan Wei
thanks: "Code and data are available at: https://github.com/wyx827/2024USpresidentialelection.git."
date: today
date-format: long
toc: true
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

#install.packages("tidymodels")
#install.packages("usmap")
library(tidyverse)
library(ggplot2)
library(rstanarm)
library(knitr)
library(usmap)

@@

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

@@

ggplot(analysis_data, aes(x = as.Date(end_date))) + 
  geom_line(stat = "count", color = "black", size = 1) + 
  labs(title = "Distribution of End Dates", x = "End Date", y = "Count") + 
  theme_minimal()
```

@@

The posterior distributions of the parameters allow us to quantify the uncertainty around each effect:\
* The coefficient for end_date informs us about how Trump’s polling percentages have evolved over time. A positive coefficient would suggest an upward trend, while a negative coefficient would indicate a decline.\
* The coefficient for numeric_grade captures the impact of pollster quality on the polling percentage. High-quality pollsters may produce different estimates compared to lower-quality ones.\
* The state-level effects account for regional differences in Trump’s support. Some states may show significantly higher or lower levels of support, even after adjusting for the time of the poll and pollster quality.

## Model Justification
@tbl-summary shows the summary table for the model. And then evaluate it on the test set. It appears as though the model is having difficulty identifying Trump supporters.
```{r}
#| echno: False
#| label: tbl-summary
#| tbl-cap: Regression Model Estimate Summary
#| echo: false
#| warning: false
#| message: false

@@

# Display the first 10 rows of the coefficients table using kable, formatted to 4 decimal places
kable(head(coefficients_table, 10), digits = 4)
```

### VIF Evaluation
@tbl-vif displays that

```{r}
#| echno: false
#| label: tbl-vif
#| tbl-cap: VIF shows a reasonable output.
#| echo: false
#| warning: false
#| message: false

library(car)
kable(vif(regression_model))
```

@@

# Define electoral votes for each state
electoral_votes <- data.frame(
  state = c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Florida", 
            "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", 
            "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", 
            "Missouri", "Montana", "Nebraska", "Nevada", 
            "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", 
            "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", 
            "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming", 
            "District of Columbia"),
  electoral_votes = c(9, 3, 11, 6, 55, 9, 7, 3, 29, 16, 4, 4, 20, 11, 6, 6, 8, 8, 4, 10, 11, 15, 10, 6, 10, 
                      3, 10, 6, 4, 14, 5, 29, 16, 3, 18, 7, 6, 20, 4, 9, 3, 11, 38, 6, 3, 13, 12, 5, 10, 3, 3)
)

# Apply the regression model to predict Trump's percentage in each state
analysis_data$predicted_percent_trump <- predict(regression_model, newdata = analysis_data)

# Summarize Trump's predicted percentage for each state
total_trump <- aggregate(predicted_percent_trump ~ state, data = analysis_data, FUN = mean)

@@

# Create a table showing the prediction results
library(knitr)

@@

@@

```{r}
#| echno: false
#| label: fig-diagnosis-1
#| tbl-cap: Linear Regression Assumptions hold
#| echo: false
#| warning: false
#| message: false

@@

## Multicolinearity Check

>>>>

---
title: "Predicting Kamala Harris's Victory in the 2024 US Election"
subtitle: "An Investigation into Poll Data Using Multiple Linear Regression"
author: 
  - Xuanle Zhou
  - Yongqi Liu
  - Yuxuan Wei
thanks: "Code and data are available at: https://github.com/wyx827/2024USpresidentialelection.git."
date: today
date-format: long
toc: true
abstract: "The study predicting the 2024 U.S. presidential election is important as it provides the predicted outcomes for the US election. Analysis is based on state-by-state electoral votes, and these outcomes can influence campaign strategies and shape public discourse. The results indicate Kamala Harris will win by over 270 votes in this election cycle."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Install the required packages
#install.packages(c("tidyverse", "ggplot2", "rstanarm", "knitr", "usmap", "arrow"))
library(tidyverse)
library(ggplot2)
library(rstanarm)
library(knitr)
library(usmap)
library(arrow)

@@

@sec-data expains the data we used to build up the prediction mode.

@@

ggplot(analysis_data, aes(x = as.Date(end_date))) + 
  geom_line(stat = "count", color = "black", size = 1) + 
  labs(title = "Distribution of End Dates", x = "End Date", y = "Count") + 
  theme_minimal()

@@

The posterior distributions of the parameters allow us to quantify the uncertainty around each effect:\
- The coefficient for end_date informs us about how Trump’s polling percentages have evolved over time. A positive coefficient would suggest an upward trend, while a negative coefficient would indicate a decline.\
- The coefficient for numeric_grade captures the impact of pollster quality on the polling percentage. High-quality pollsters may produce different estimates compared to lower-quality ones.\
- The state-level effects account for regional differences in Trump’s support. Some states may show significantly higher or lower levels of support, even after adjusting for the time of the poll and pollster quality.

## Model Justification
@tbl-summary shows the summary table for the model. And then evaluate it on the test set. It appears as though the model is having difficulty identifying Trump supporters. The reson why the multiple regression is applied is because.....

```{r}
#| echno: false
#| label: tbl-summary
#| tbl-cap: Regression Model Estimate Summary
#| echo: false
#| warning: false
#| message: false

@@

# Display the first 10 rows of the coefficients table using kable, formatted to 4 decimal places
kable(head(coefficients_table, 10), digits = 2)
```

@@

# Read the test data from the .parquet file
analysis_data_test <- read_parquet(here::here("data/02-analysis_data/test_data.parquet"))
analysis_data_train<- read_parquet(here::here("data/02-analysis_data/train_data.parquet"))
# Filter out states in the test data that don't exist in the training data

analysis_data_test <- analysis_data_test %>%
 filter(state %in% unique(analysis_data_train$state))

# Define electoral votes for each state
electoral_votes <- data.frame(
  state = c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Florida",
            "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", 
            "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", 
            "Missouri", "Montana", "Nebraska", "Nevada", 
            "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", 
            "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", 
            "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming", 
            "District of Columbia"),
  electoral_votes = c(9, 3, 11, 6, 55, 9, 7, 3, 29, 16, 4, 4, 20, 11, 6, 6, 8, 8, 4, 10, 11, 15, 10, 6, 10, 
                      3, 10, 6, 4, 14, 5, 29, 16, 3, 18, 7, 6, 20, 4, 9, 3, 11, 38, 6, 3, 13, 12, 5, 10, 3, 3))

# Apply the regression model to predict Trump's percentage in each state
analysis_data_test$predicted_percent_trump <- predict(regression_model, newdata = analysis_data_test)

# Summarize Trump's predicted percentage for each state
total_trump <- aggregate(predicted_percent_trump ~ state, data = analysis_data_test, FUN = mean)

@@

@@

# Replace NA values in the 'winner' column with 'Trump'
#state_predictions$winner[is.na(state_predictions$winner)] <- "Trump"

@@

```{r}
#| echo: false
#| label: fig-diagnosis-3
#| tbl-cap: Linear Regression Assumptions hold
#| warning: false
#| message: false
#| fig.width: 10
#| fig.height: 8

@@

## Calculate Mean Squared Error (MSE) on test data
```{r}
#| echo: false
#| label: fig-diagnosis-1
#| tbl-cap: Calculate Mean Squared Error (MSE) on test data
#| warning: false
#| message: false

predicted_values <- predict(regression_model, newdata = analysis_data_test)
actual_values <- analysis_data_test$percent

mse <- mean((predicted_values - actual_values)^2)
print(mse)
```

```{r}

library(caret)
set.seed(856)
train_control <- trainControl(method = "cv", number = 10)
cv_model <- train(percent ~ numeric_grade + sample_size + state + transparency_score + end_date,
                  data = analysis_data_train,
                  method = "lm",
                  trControl = train_control)

print(cv_model$results)

```{r}
baseline_prediction <- mean(analysis_data_train$percent)
baseline_mse <- mean((analysis_data_test$percent - baseline_prediction)^2)
print(baseline_mse)
```

```{r}
# Calculate R-squared on test data
rss <- sum((predicted_values - actual_values)^2)  # Residual Sum of Squares
tss <- sum((actual_values - mean(actual_values))^2)  # Total Sum of Squares

r_squared <- 1 - (rss/tss)
print(r_squared)
```

```{r}
# Create a correlation matrix to check for collinearity
cor_matrix <- cor(analysis_data_train[, c("numeric_grade", "sample_size", "transparency_score")], use = "complete.obs")
print(cor_matrix)
```
```{r}
# Check for sparse levels in 'state'
state_counts <- table(analysis_data_train$state)
print(state_counts)
```
